---
title: '`baselinenowcast` Analysis Pipeline'
author: "Kaitlyn Johnson"
date: "2025-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Introduction

This document defines the analysis pipeline used to evaluate the performance
of the `baselinenowcast` R package, which is intended to be used as a
baseline method for nowcasting right-truncated epidemiological data.
More details on the method can be found in the `baselinenowcast` repository.

The performance of the baseline nowcast method will be evaluated as follows:
1. Applied to real-world data from the German Nowcast Hub from 1 Dec 2021 to
1 April 2022, nationally, by age group, and by state.
2. Applied to 4 simulated data scenarios in which we take two age strata, one
sparse and one rich from the German Nowcast data, and apply a long and short
delay to the data. For each data scenario, we will assess 24 different
combinations of model specifications
3. Applied to real-world data from 1. 2013-2014 Measles outbreak in the
Netherlands 2. 2023-2024 Norovirus outbreak in the UK and 3. Comparison of
performance of the mean/median ensemble in the German Nowcast Hub

This workflow uses the `targets` package.

```{r, eval = FALSE}
rmarkdown::render("_targets.Rmd")
```

The pipeline can be run using,
```{r, eval = FALSE}
tar_make()
```

# Setup

Set up the workflow pipeline.
```{r}
library(targets)
library(crew)
library(RcppTOML)
library(readr)

controller <- crew_controller_local(
  workers = 8,
  seconds_idle = 600
)
```

Define the shared global options and load the R functions from the `R` folder
```{targets globals tar_globals = TRUE}
library(targets)
library(ggplot2)
library(purrr, quietly = TRUE)
library(arrow)
library(here)
functions <- list.files(here("R"), full.names = TRUE)
walk(functions, source)
rm("functions")

tar_option_set(
  packages = c(
    "tibble", "dplyr", "lubridate",
    "baselinenowcast",
    "readr"
  ),
  workspace_on_error = TRUE,
  # Run with a pre-specified crew controller
  controller = controller,
  # Setup storage on workers vs on the main node.
  memory = "transient",
  storage = "worker",
  retrieval = "worker",
  format = "parquet", # default storage format
  error = NULL # tells errored targets to return NULL rather than
  # have whole pipeline fail
)
```
