---
title: "Summary metrics for paper"
author: "Kaitlyn Johnson"
date: "2025-06-16"
output: html_document
---

```{r, echo = FALSE}
library(targets)
library(dplyr)
library(kableExtra)
library(scoringutils)
library(tidyr)

# Hub validation outputs
validation_scores <- tar_read(validation_scores)
validation_coverage <- tar_read(validation_coverage)
mean_delay_validation <- tar_read(mean_delay_validation)

# Model permutation outputs
scores_mp <- tar_read(scores_mp)
coverage_mp <- tar_read(coverage_mp)

# Norovirus outputs
scores_noro <- tar_read(noro_scores)
coverage_noro <- tar_read(noro_coverage)
```

# Hub validation results

Overall relative WIS: age groups
```{r, echo = FALSE}
scores_sum <- validation_scores |>
  filter(age_group != "00+") |>
  summarise_scores(by = "model")

comparison_score_rt <- scores_sum |>
  filter(model == "original KIT simple nowcast") |>
  rename(KIT_rt_wis = wis) |>
  select(KIT_rt_wis)

comparison_score_rev <- scores_sum |>
  filter(model == "revised KIT simple nowcast") |>
  rename(KIT_rev_wis = wis) |>
  select(KIT_rev_wis)

rel_wis <- scores_sum |>
  filter(model == "baselinenowcast") |>
  bind_cols(comparison_score_rt) |>
  mutate(
    relative_wis = wis / pmax(KIT_rt_wis, .Machine$double.eps),
    strata = "age groups",
    relative_to = "original KIT simple nowcast"
  ) |>
  select(relative_wis, relative_to, model)

rel_wis_rev <- scores_sum |>
  filter(model == "baselinenowcast") |>
  bind_cols(comparison_score_rev) |>
  mutate(
    relative_wis = wis / pmax(KIT_rev_wis, .Machine$double.eps),
    strata = "age groups",
    relative_to = "revised KIT simple nowcast"
  ) |>
  select(relative_wis, relative_to, model)
```
Percent contribution to WIS by age group
```{r, echo = FALSE}
validation_scores |>
  filter(age_group != "00+") |>
  filter(model == "baselinenowcast") |>
  summarise_scores(by = "age_group") |>
  mutate(
    total_wis = sum(wis),
    pct_contribution = wis / total_wis,
    mean_wis = mean(wis)
  ) |>
  kable(caption = "WIS percent contribution by age group")
```
Overall relative WIS: national
```{r, echo = FALSE}
scores_sum <- validation_scores |>
  filter(age_group == "00+") |>
  summarise_scores(by = "model")

comparison_score_rt <- scores_sum |>
  filter(model == "original KIT simple nowcast") |>
  rename(KIT_rt_wis = wis) |>
  select(KIT_rt_wis)

comparison_score_rev <- scores_sum |>
  filter(model == "revised KIT simple nowcast") |>
  rename(KIT_rev_wis = wis) |>
  select(KIT_rev_wis)

rel_wis_ntl <- scores_sum |>
  filter(model == "baselinenowcast") |>
  bind_cols(comparison_score_rt) |>
  mutate(
    relative_wis = wis / pmax(KIT_rt_wis, .Machine$double.eps),
    strata = "national",
    relative_to = "original KIT simple nowcast"
  ) |>
  select(relative_wis, relative_to, model)

rel_wis_rev_ntl <- scores_sum |>
  filter(model == "baselinenowcast") |>
  bind_cols(comparison_score_rev) |>
  mutate(
    relative_wis = wis / pmax(KIT_rev_wis, .Machine$double.eps),
    strata = "national",
    relative_to = "revised KIT simple nowcast"
  ) |>
  select(relative_wis, relative_to, model)
```

Join together results and print
```{r, echo = FALSE}
bind_rows(rel_wis, rel_wis_rev) |>
  kable(caption = "Overall relative WIS by age group")
bind_rows(rel_wis_ntl, rel_wis_rev_ntl) |>
  kable(caption = "Overall relative WIS national")
```

Overall score breakdown
```{r, echo = FALSE}
validation_scores |>
  filter(age_group != "00+") |>
  group_by(model) |>
  summarise(
    pct_under = sum(underprediction) / sum(wis),
    pct_over = sum(overprediction) / sum(wis),
    pct_disp = sum(dispersion) / sum(wis)
  ) |>
  kable(caption = "WIS percent breakdown by model for age groups")
```

Score summary by age group
```{r, echo = FALSE}
validation_scores |>
  filter(age_group != "00+") |>
  summarise_scores(by = c("model", "age_group")) |>
  group_by(model) |>
  summarise(
    min_wis = min(wis),
    age_group_min_wis = age_group[which.min(wis)],
    max_wis = max(wis),
    age_group_max_wis = age_group[which.max(wis)]
  ) |>
  kable(caption = "Min and max WIS by age group and model")
```

Mean delay by age
```{r, echo = FALSE}
mean_delay_validation |>
  group_by(age_group) |>
  summarise(mean_delay = sum(delay_time * mean_delay)) |>
  ungroup() |>
  summarise(
    min_delay = min(mean_delay),
    min_delay_age_group = age_group[which.min(mean_delay)],
    max_delay = max(mean_delay),
    max_delay_age_group = age_group[which.max(mean_delay)]
  ) |>
  kable(caption = "Min and max mean delay by age group")
```

Relative WIS by age group
```{r, echo = FALSE}
scores_sum <- validation_scores |>
  filter(age_group != "00+") |>
  summarise_scores(by = c("model", "age_group"))

comparison_score_rt <- scores_sum |>
  filter(model == "KIT simple nowcast revised") |>
  rename(KIT_rt_wis = wis) |>
  select(KIT_rt_wis, age_group)

rel_wis <- scores_sum |>
  filter(model == "baselinenowcast") |>
  left_join(comparison_score_rt, by = "age_group") |>
  mutate(
    relative_wis = wis / pmax(KIT_rt_wis, .Machine$double.eps),
    strata = "age groups",
    relative_to = "KIT simple nowcast original"
  ) |>
  select(relative_wis, relative_to, age_group, model)

rel_wis |>
  summarise(
    min_rel_wis = min(relative_wis),
    min_rel_wis_age_group = age_group[which.min(relative_wis)],
    max_rel_wis = max(relative_wis),
    max_rel_wis_age_group = age_group[which.max(relative_wis)]
  ) |>
  kable(caption = "Min and max relative WIS by age group")
```

Coverage by model
```{r}
validation_coverage |>
  group_by(model, interval_range) |>
  summarise(empirical_coverage = sum(interval_coverage) / n()) |>
  filter(interval_range %in% c(50, 95)) |>
  pivot_wider(
    names_from = interval_range,
    values_from = empirical_coverage
  ) |>
  kable(caption = "Model validation: coverage summary")
```

# Model permutation results
```{r, echo = FALSE}
scores_mp |>
  filter(age_group != "00+") |>
  summarise_scores(by = "model_variation_string") |>
  kable(caption = "Model permutations: overall scores")
```
Model permutation relative WIS:
```{r, echo = FALSE}
sum_scores <- scores_mp |>
  filter(age_group != "00+") |>
  summarise_scores(by = "model_variation_string")

default_scores <- sum_scores |>
  filter(model_variation_string == "Default") |>
  mutate(default_wis = wis) |>
  select(default_wis)
default_wis <- default_scores$default_wis[1]

sum_scores |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis) |>
  kable(caption = "Model permutations: relative scores")
```
Model permutation relative WIS by age group:
```{r, echo = FALSE}
sum_scores <- scores_mp |>
  filter(age_group != "00+") |>
  summarise_scores(by = c("model_variation_string", "age_group"))

default_scores <- sum_scores |>
  filter(model_variation_string == "Default") |>
  mutate(default_wis = wis) |>
  select(default_wis, age_group)

sum_scores |>
  left_join(default_scores, by = "age_group") |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis, age_group) |>
  filter(model_variation_string == "Borrowed estimates from all age groups") |>
  kable(caption = "Model permutation Borrowing: relative scores by age group")

sum_scores |>
  left_join(default_scores, by = "age_group") |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis, age_group) |>
  filter(!model_variation_string %in% c(
    "Borrowed estimates from all age groups",
    "Complete reporting triangle",
    "Default"
  )) |>
  kable(caption = "Model permutations training volume: relative scores by age group")
sum_scores |>
  left_join(default_scores, by = "age_group") |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis, age_group) |>
  filter(!model_variation_string %in% c(
    "Borrowed estimates from all age groups",
    "Complete reporting triangle",
    "Default"
  )) |>
  summarise(
    min_rel_wis = min(rel_wis),
    age_group_min = age_group[which.min(rel_wis)],
    model_var_rel_min = model_variation_string[which.min(rel_wis)],
    max_rel_wis = max(rel_wis),
    age_group_max = age_group[which.max(rel_wis)],
    model_var_rel_max = model_variation_string[which.max(rel_wis)]
  ) |>
  kable(caption = "Model permutations: training volumes min and max scores by age group")



sum_scores |>
  left_join(default_scores, by = "age_group") |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis, age_group) |>
  filter(model_variation_string == "Complete reporting triangle") |>
  kable(caption = "Model permutations complete rep tri: relative scores by age group")

sum_scores |>
  left_join(default_scores, by = "age_group") |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis, age_group) |>
  filter(model_variation_string == "Complete reporting triangle") |>
  summarise(
    min_rel_wis = min(rel_wis),
    age_group_min = age_group[which.min(rel_wis)],
    model_var_rel_min = model_variation_string[which.min(rel_wis)],
    max_rel_wis = max(rel_wis),
    age_group_max = age_group[which.max(rel_wis)],
    model_var_rel_max = model_variation_string[which.max(rel_wis)]
  ) |>
  kable(caption = "Model permutations: complete reporting triangle min and max scores by age group")
# Borrowing
sum_scores |>
  left_join(default_scores, by = "age_group") |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis, age_group) |>
  filter(model_variation_string == "Borrowed estimates from all age groups") |>
  kable(caption = "Model permutations Borrowing: relative scores by age group")

sum_scores |>
  left_join(default_scores, by = "age_group") |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis, age_group) |>
  filter(model_variation_string == "Complete reporting triangle") |>
  summarise(
    min_rel_wis = min(rel_wis),
    age_group_min = age_group[which.min(rel_wis)],
    model_var_rel_min = model_variation_string[which.min(rel_wis)],
    max_rel_wis = max(rel_wis),
    age_group_max = age_group[which.max(rel_wis)],
    model_var_rel_max = model_variation_string[which.max(rel_wis)]
  ) |>
  kable(caption = "Model permutations: complete reporting triangle min and max scores by age group")
```

Model permutations horizon
```{r, echo = FALSE}
sum_scores <- scores_mp |>
  filter(age_group != "00+") |>
  summarise_scores(by = c("model_variation_string", "horizon"))

default_scores <- sum_scores |>
  filter(model_variation_string == "Default") |>
  mutate(default_wis = wis) |>
  select(default_wis, horizon)

sum_scores |>
  left_join(default_scores, by = "horizon") |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis, horizon) |>
  filter(model_variation_string == "Borrowed estimates from all age groups") |>
  kable(caption = "Model permutation Borrowing: relative scores by horizon")

sum_scores |>
  left_join(default_scores, by = "horizon") |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis, horizon) |>
  filter(!model_variation_string %in% c(
    "Borrowed estimates from all age groups",
    "Complete reporting triangle",
    "Baseline validation approach"
  )) |>
  kable(caption = "Model permutations training volume: relative scores by horizon")
sum_scores |>
  left_join(default_scores, by = "horizon") |>
  mutate(rel_wis = wis / default_wis) |>
  select(model_variation_string, rel_wis, horizon) |>
  filter(!model_variation_string %in% c(
    "Borrowed estimates from all age groups",
    "Complete reporting triangle",
    "Baseline validation approach"
  )) |>
  summarise(
    min_rel_wis = min(rel_wis),
    horizon_min = horizon[which.min(rel_wis)],
    model_var_rel_min = model_variation_string[which.min(rel_wis)],
    max_rel_wis = max(rel_wis),
    horizon_max = horizon[which.max(rel_wis)],
    model_var_rel_max = model_variation_string[which.max(rel_wis)]
  ) |>
  kable(caption = "Model permutations: training volumes min and max scores by horizon")
```
Coverage summary
```{r, echo = FALSE}
coverage_mp |>
  group_by(model, interval_range, model_variation_string) |>
  summarise(empirical_coverage = sum(interval_coverage) / n()) |>
  pivot_wider(
    names_from = interval_range,
    values_from = empirical_coverage
  ) |>
  kable(caption = "Model permutations: coverage summary")
```

# Norovirus results

Scores compared to Mellor et al baseline
```{r, echo = FALSE}
scores_sum <- scores_noro |>
  scoringutils::summarise_scores(by = "model") |>
  select(model, wis)

baseline_comparison <- scores_sum |>
  filter(model == "baseline Mellor et al") |>
  rename(comparison_wis = wis) |>
  select(comparison_wis) |>
  pull()

scores_sum |>
  filter(model != "baseline Mellor et al") |>
  mutate(comparison_wis = baseline_comparison) |>
  mutate(rel_wis = wis / pmax(comparison_wis, .Machine$double.eps)) |>
  kable(caption = "Relative WIS compared to baseline Mellor et al")
```
Scores compared to baselnenowcast default
```{r, echo = FALSE}
scores_sum <- scores_noro |>
  scoringutils::summarise_scores(by = "model") |>
  select(model, wis)

baseline_comparison <- scores_sum |>
  filter(model == "baselinenowcast base") |>
  rename(comparison_wis = wis) |>
  select(comparison_wis) |>
  pull()

scores_sum |>
  filter(model != "baselinenowcast base") |>
  mutate(comparison_wis = baseline_comparison) |>
  mutate(rel_wis = wis / pmax(comparison_wis, .Machine$double.eps)) |>
  kable(caption = "Relative WIS compared to baselinenowcast base")
```


Scores compared to epinowcast
```{r, echo = FALSE}
scores_sum <- scores_noro |>
  scoringutils::summarise_scores(by = "model") |>
  select(model, wis)

baseline_comparison <- scores_sum |>
  filter(model == "epinowcast") |>
  rename(comparison_wis = wis) |>
  select(comparison_wis) |>
  pull()

scores_sum |>
  filter(model != "epinowcast") |>
  mutate(comparison_wis = baseline_comparison) |>
  mutate(rel_wis = wis / pmax(comparison_wis, .Machine$double.eps)) |>
  kable(caption = "Relative WIS compared to epinowcast Mellor et al")
```
Scores compared to GAM

```{r, echo = FALSE}
scores_sum <- scores_noro |>
  scoringutils::summarise_scores(by = "model") |>
  select(model, wis)

baseline_comparison <- scores_sum |>
  filter(model == "GAM") |>
  rename(comparison_wis = wis) |>
  select(comparison_wis) |>
  pull()

scores_sum |>
  filter(model != "GAM") |>
  mutate(comparison_wis = baseline_comparison) |>
  mutate(rel_wis = wis / pmax(comparison_wis, .Machine$double.eps)) |>
  kable(caption = "Relative WIS compared to GAM Mellor et al")
```

Coverage summary
```{r, echo = FALSE}
coverage_noro |>
  group_by(model, interval_range) |>
  summarise(empirical_coverage = sum(interval_coverage) / n()) |>
  pivot_wider(
    names_from = interval_range,
    values_from = empirical_coverage
  ) |>
  kable(caption = "norovirus: coverage summary")
```
